# Homework 1: Using R for Distances and Mapping
Homework 1 (lecture 1 slide 31)
Spatial Statistics
Daniel J. Hocking
Started: 04 September 2013

## P0: Read Sections 3.1 and 3.2 of the text

## P1: Find the geo-coordinates (longitude, latitude) of New York, Seattle, and Los Angeles and calculate the pairwise distances:

```{r}
# Problem 1
library(maps)
library(mapproj)
library(knitr)

# From Google
# NYC = 40.6700° N, 73.9400° W
# LA = 34.0500° N, 118.2500° W

NYC.lat <- 40.6700
NYC.lon <- -73.9400
LA.lat <- 34.0500
LA.lon <- -118.2500

# Convert to Radians
NYC.lat.rad <- NYC.lat*pi/180
NYC.lon.rad <- NYC.lon*pi/180
LA.lat.rad <- LA.lat*pi/180
LA.lon.rad <- LA.lon*pi/180
```

### a) naively: by using Euclidean distances directly on the longitude/latitude
Use Pythagorean theorem to solve
```{r}
dist.euclid.rad <- sqrt((NYC.lat.rad - LA.lat.rad)^2 + (NYC.lon.rad - LA.lon.rad)^2) 
dist.euclid <- sqrt((NYC.lat - LA.lat)^2 + (NYC.lon - LA.lon)^2) 
```

The Euclidean distance = 44.8018 degrees. Given the lack of projection, it is impossible to correctly convert this to relevant distances such as kilometers. In radians the distance = 0.78194 radians.

### b) correctly by using the geodesic distances
```{r}
dist.geo <- 3678*acos(sin(NYC.lat.rad)*sin(LA.lat.rad) + cos(NYC.lat.rad)*cos(LA.lat.rad)*cos(NYC.lon.rad - LA.lon.rad))
```

The geodesic distance between NYC and LA is 2,276 km. For comparison with the previous answer, this can be divided by 3678 to convert to radians `dist.geo/3678` = 0.6188 radians.

### c) by using straight line connections that cut through the earth (apply some geometry here)


## P2: Obtain a map of the US states and project it using the three popular
projections: Albers, Lambert and Bonne. Comment on the resulting graphs. Use:
```{r}
    library(maps); library(mapproj)
    par(mfrow=c(2,2), mar=c(0,0,2,0))   # puts multiple graphs on page
    map("state"); title("long and lat")
    map("state",proj='albers', par=c(30,40)); title("albers")
    map("state",proj='bonne', param=35); title("bonne")
    map("state",proj='lambert', par=c(30,40)); title("lambert")
```

The longitude and latitude projection changes the orientation of the states by stretching the top of the map more than the bottom compared with the other plots. The western border with Canada is also horizontal. The result is that Maine and Oregon appear much farther from each other and the states are oriented differently. At the size of the plots on my screen, the Albers, Bonne, and Lambert projections all apprea very similar.

## P3: Obtain a map that contains Greenland and the contiguous USA. Plot a long/lat map and then choose 3 projections – select appropriate parameters for these. Comment on the effect of the projections. R code for unprojected map:
```{r}
par(mfrow = c(1,1))
map("world", xlim=c(-130, -20),ylim=c(25,85)); title("long,lat") # map.axes(); # optional: adds axes to the map!
```


